USCIS data on status adjustment from non-immigrant visa status to US permanent residency
========================================================
Analysis - Countries of birth and US non-immigrant visa types
By:       Sasi Jonnalagadda  
Date:     `r date()`

Key questions to be addressed:
1. What is the timing of the civil surgeon examination, a mandatory medical exam for application for status adjustment, since the time of entry into the US as a non-immigrant?
2. What is the impact of this exam and it's timing on prevention of TB cases among foreign born individuals in the US?

Key questions addressed in this analysis:
1. Which countries do applicants for status adjustment represent most frequently?
2. What non-immigrant visa types fo these applicants enter the US on?
3. How long are the non-immigrants staying in the US before adjusting their status?  The dataset from USCIS does not contain dates on the civil surgeon exam which happens approximately 1 year before or after application for status adjustment is filed.  Therefore, the application date in the USCIS dataset has been used as a proxy for the civil surgeon exam.

```{r options, echo=FALSE}
# Set options for report
opts_chunk$set(comment = NA)
opts_chunk$set(echo = FALSE)
opts_chunk$set(message = FALSE)
```

Dataset being used for this is ni.main
```{r "dataset",cache=TRUE}
load("~/Analyses/uscis data/uscis/uscis/ni.main.Rda")
str(ni.main)
```

#Summary of time to exam (variable - texamyr)

```{r "summary"}
summary(ni.main$texamyr)
require(ggplot2)
qplot(texamyr, data=ni.main, geom = "histogram", binwidth=0.1, xlim=c(0,20)) +
  xlab("Time from last arrival in US to application for status adjustment (year)") + 
  ylab("Frequency") + opts(title="Non-immigrant applicants for status adjustment")
```

1. Most common countries of birth of applicants for status adjustment

*1a. By WHO defined TB regions (TB incidence rate of <20, 20-100, >100/100,000)

``` {r "country of birth",cache=TRUE}
#Most common countries - this list was generated in the datasets2 code 
#includes countries with >20,000 individuals per country in the uscis dataset

#Aggregate over COB (country of birth) and tbburden (3 categories of TB incidence (<20, 20-100, >100/100,000) )
aggdata <- aggregate(ni.main, by=list(ni.main$COB, ni.main$tbburden), FUN=length)
keepvars <- names(aggdata) %in% c("Group.1", "Group.2", "COB")
aggdata <- aggdata[keepvars]
###Rename variables
require(reshape)
aggdata <- rename(aggdata, c(Group.1="cob", Group.2="tbburden", COB="Number"))
aggdata <- aggdata[order(-aggdata$Number),]
list(aggdata)
```

*1b. Plotting bargraphs by country of birth

```{r "country of birth (plot)",fig.height=8, fig.width=8, fig.align='center'}
#Plotting bargraphs
require(ggplot2)
require(scales)

###Code to make the graph look ordered by frequency and not alphabetical order of country of birth (cob)
cob_order <- transform(aggdata, cob=reorder(cob, Number)) #Countries appreas from low to high frequency

#this plot will contain cob on x-axis
p <- ggplot(data=cob_order, aes(x=cob, y=Number)) + geom_bar(colour="grey", width=.8) + 
  xlab("Country of birth") + ylab("Frequency") + opts(title="Country of birth of non-immigrant applicants for status adjustment")

#this plot will flip the axes and show cob on y-axis (easier to read)
p + geom_bar() + coord_flip() + scale_y_continuous(labels=comma, breaks=c(20000, 100000, 200000, 300000, 400000, 500000, 600000))
```

*1c. Country of birth stratified by TB regions

```{r "country of birth by TB regions", fig.height=5, fig.width=15, fig.align='center'}
#By WHO TB regions (Facet by tbburden (WHO))

p + geom_bar() + theme_bw() + coord_flip() +
  scale_y_continuous(labels=comma, breaks=c(20000, 100000, 200000, 300000, 400000, 500000, 600000)) +
  facet_wrap(~ tbburden, ncol=3, scales="free") +
  opts(title="Country of birth of non-immigrant applicants for status adjustment, by TB burden")
```

*1d. By Yecai defined TB regions (TB incidence rate of <15, 15-100, >100/100,000)

``` {r "Yecai TB classification", fig.height=5, fig.width=15, fig.align='center',cache=TRUE}
#Aggregate over COB and tbburden.yecai
aggdata.yecai <- aggregate(ni.main, by=list(ni.main$COB, ni.main$tbburden.yecai), FUN=length)
keepvars <- names(aggdata.yecai) %in% c("Group.1", "Group.2", "COB")
aggdata.yecai <- aggdata.yecai[keepvars]
###Rename variables
require(reshape)
aggdata.yecai <- rename(aggdata.yecai, c(Group.1="cob", Group.2="tbburden.yecai", COB="Freq"))

##By Yecai TB regions
require(ggplot2)
require(scales)

###Code to make the graph look ordered by frequency and not alphabetical order of country of birth (cob)
cob_order <- transform(aggdata.yecai, cob=reorder(cob, Freq)) #Countries appreas from low to high frequency

#this plot will contain cob on x-axis

p <- ggplot(data=cob_order, aes(x=cob, y=Freq)) + geom_bar(colour="grey", width=.8) + 
  xlab("Country of birth") + ylab("Frequency") + opts(title="Country of birth of non-immigrant applicants for status adjustment")

p + geom_bar() + theme_bw() + coord_flip() +
  scale_y_continuous(labels=comma, breaks=c(20000, 100000, 200000, 300000, 400000, 500000, 600000)) +
  facet_wrap(~ tbburden.yecai, ncol=3, scales="free") +
  opts(title="Country of birth of non-immigrant applicants for status adjustment, by TB burden (Yecai classification)")
```


2. Most common visa types 

```{r "visa types", fig.align='center',cache=TRUE}
#Most common visa types
require(ggplot2)
require(reshape)
library(scales)
visacollapse <- as.data.frame(table(droplevels(ni.main)$visacat)) #This takes the table command and converts the output into a dataframe which can then be analyzed
visacollapse <- visacollapse[order(-visacollapse$Freq),] #order the freq of visatypes from high to low
visacollapse <- rename(visacollapse, c(Var1="visa"))

###Pre-graphing code for ordering the visa types by frequency and not their numeric code 
visacollapse_reorder <- transform(visacollapse, visa=reorder(visa, Freq)) #makes graph appear with visa types from low to high freq

###This keep the visa type on the x-axis
pvisatype <- ggplot(data=visacollapse_reorder, aes(x=visa, y=Freq)) + geom_bar(colour="grey", width=.8) + 
  xlab("Visa Type") + ylab("Frequency") + opts(title="Non-immigrant applicants for status adjustment")

###This flips the coordinates (coord_flip) 
pvisatype + geom_bar() + coord_flip() + scale_y_continuous(labels=comma, breaks=c(25000, 100000, 200000, 300000, 400000, 500000, 600000, 700000))
```

*2a. Visa categories defined by Yecai

```{r "Yecai's visa categories", fig.height=5, fig.width=15, fig.align='center',cache=TRUE}
##Yecai's visatypes
visacollapse.yecai <- as.data.frame(table(droplevels(ni.main)$visacat.yecai)) #This takes the table command and converts the output into a dataframe which can then be analyzed
require(reshape)
visacollapse.yecai <- rename(visacollapse.yecai, c(Var1="visa"))

###Pre-graphing code for ordering the visa types by frequency and not their numeric code 
visacollapse.yecai_reorder <- transform(visacollapse.yecai, visa=reorder(visa, Freq)) #makes graph appear with visa types from low to high freq

###This keep the visa type on the x-axis
require(ggplot2)
pvisatype <- ggplot(data=visacollapse.yecai_reorder, aes(x=visa, y=Freq)) + geom_bar(colour="grey", width=.3) + 
  xlab("Visa Type") + ylab("Frequency") + opts(title="Non-immigrant applicants for status adjustment, by Yecai's visa classification")

###This flips the coordinates (coord_flip) 
require(scales)
pvisatype + geom_bar(colour="grey", width=.5) + coord_flip() + scale_y_continuous(labels=comma, breaks=c(25000, 100000, 200000, 300000, 400000, 500000, 600000, 700000, 800000, 900000))
```


