USCIS data on status adjustment from non-immigrant visa status to US permanent residency
========================================================
Analysis - Time to status adjustment, by visa types stratified by WHO TB region, geographical region and country of birth
By:       Sasi R Jonnalagadda, phd  
Date:     `r date()`

#This report is for Asian countries

```{r options, echo=FALSE}
# Set options for report
opts_chunk$set(comment = NA)
opts_chunk$set(echo = FALSE)
opts_chunk$set(message = FALSE)
```

Dataset being used for this is ni_main
```{r "dataset",cache=TRUE}
load("~/Analyses/uscis data/uscis/uscis/ni.main.Rda")
str(ni.main)
```

```{r "defining survival frame"}
# define custom function to create a survival data.frame
#This data.frame will allow use of ggplot to plot the survival function (rather than using plot)
#The original code for creating this data.frame (provide url) was modified to plot the hazard function (cumevent) rather than the survival function
createSurvivalFrame <- function(f.survfit){
  # initialise frame variable
  f.frame <- NULL
  # check if more then one strata
  if(length(names(f.survfit$strata)) == 0){
    # create data.frame with data from survfit
    f.frame <- data.frame(time=f.survfit$time, n.risk=f.survfit$n.risk, n.event=f.survfit$n.event, n.censor = f.survfit
                          $n.censor, surv=f.survfit$surv, cumevent=f.survfit$cumevent, upper=f.survfit$upper, lower=f.survfit$lower)
    # create first two rows (start at 1)
    f.start <- data.frame(time=c(0, f.frame$time[1]), n.risk=c(f.survfit$n, f.survfit$n), n.event=c(0,0),
                          n.censor=c(0,0), surv=c(1,1), cumevent=c(0,0), upper=c(1,1), lower=c(1,1))
    # add first row to dataset
    f.frame <- rbind(f.start, f.frame)
    # remove temporary data
    rm(f.start)
  }
  else {
    
    # create vector for strata identification
    f.strata <- NULL
    for(f.i in 1:length(f.survfit$strata)){
      # add vector for one strata according to number of rows of strata
      f.strata <- c(f.strata, rep(names(f.survfit$strata)[f.i], f.survfit$strata[f.i]))
    }
    # create data.frame with data from survfit (create column for strata)
    f.frame <- data.frame(time=f.survfit$time, n.risk=f.survfit$n.risk, n.event=f.survfit$n.event, n.censor = f.survfit
                          $n.censor, surv=f.survfit$surv, cumevent=f.survfit$cumevent, upper=f.survfit$upper, lower=f.survfit$lower, strata=factor(f.strata))
    # remove temporary data
    rm(f.strata)
    # create first two rows (start at 1) for each strata
    for(f.i in 1:length(f.survfit$strata)){
      # take only subset for this strata from data
      f.subset <- subset(f.frame, strata==names(f.survfit$strata)[f.i])
      # create first two rows (time: 0, time of first event)
      f.start <- data.frame(time=c(0, f.subset$time[1]), n.risk=rep(f.survfit[f.i]$n, 2), n.event=c(0,0),
                            n.censor=c(0,0), surv=c(1,1), cumevent=c(0,0), upper=c(1,1), lower=c(1,1), strata=rep(names(f.survfit$strata)[f.i],
                                                                                                                  2))
      # add first two rows to dataset
      f.frame <- rbind(f.start, f.frame)
      # remove temporary data
      rm(f.start, f.subset)
    }
    # reorder data
    f.frame <- f.frame[order(f.frame$strata, f.frame$time), ]
    # rename row.names
    rownames(f.frame) <- NULL
  }
  # return frame
  return(f.frame)
}

# define custom function to draw kaplan-meier curve with ggplot
qplot_survival <- function(f.frame, f.CI="default", f.shape=3){
  # use different plotting commands dependig whether or not strata's are given
  if("strata" %in% names(f.frame) == FALSE){
    # confidence intervals are drawn if not specified otherwise
    if(f.CI=="default" | f.CI==TRUE ){
      # create plot with 4 layers (first 3 layers only events, last layer only censored)
      # hint: censoring data for multiple censoring events at timepoint are overplotted
      # (unlike in plot.survfit in survival package)
      ggplot(data=f.frame) + geom_step(aes(x=time, y=cumevent), direction="hv") + geom_step(aes(x=time,
                                                                                                y=upper), directions="hv", linetype=2) + geom_step(aes(x=time,y=lower), direction="hv", linetype=2)
    }
    else {
      # create plot without confidence intervalls
      ggplot(data=f.frame) + geom_step(aes(x=time, y=cumevent), direction="hv")
    }
  }
  else {
    if(f.CI=="default" | f.CI==FALSE){
      # without CI
      ggplot(data=f.frame, aes(group=strata, colour=strata)) + geom_step(aes(x=time, y=cumevent),
                                                                         direction="hv")
    }
    else {
      # with CI (hint: use alpha for CI)
      ggplot(data=f.frame, aes(colour=strata, group=strata)) + geom_step(aes(x=time, y=cumevent),
                                                                         direction="hv") + geom_step(aes(x=time, y=upper), directions="hv", linetype=2, alpha=0.5) +
                                                                           geom_step(aes(x=time,y=lower), direction="hv", linetype=2, alpha=0.5)
    }
  }
}
```

```{r "libraries"}
#create a new variable status for event=application for status adjustment; everyone gets a 1 in this dataset - no censoring for other reasons.
require(survival)
require(ggplot2)
require(scales)
require(plyr)
```


4. Time to status adjustment for different visa types-by country and region

##Data stratified by WHO TB burden, geographical region and country of birth

a. Countries with high TB incidence (>100/100,000)
```{r "Data by TB burden", cache=TRUE}
require(plyr)
require(xtable)

hightb <- subset(ni.main, tbburden=="high (TB incidence>100/100,000)", drop=TRUE) 
hightb$tbburden <- factor(hightb$tbburden)
hightb$COB <- factor(hightb$COB)

mediumtb <- subset(ni.main, tbburden=="medium (TB incidence 20-100/100,000)", drop=TRUE) 
mediumtb$tbburden <- factor(mediumtb$tbburden)
mediumtb$COB <- factor(mediumtb$COB)

lowtb <- subset(ni.main, tbburden=="low (TB incidence<20/100,000)", drop=TRUE) 
lowtb$tbburden <- factor(lowtb$tbburden)
lowtb$COB <- factor(lowtb$COB)
```

b. Defining geographical region for countries, by WHO TB burden

```{r "Data by TB burden and geographical region - Asia",cache=TRUE}
#Asia
asiahightb <- subset(hightb, region=="Asia", drop=TRUE) 
asiahightb$region <- factor(asiahightb$region)
asiahightb$COB <- factor(asiahightb$COB)

asiamedtb <- subset(mediumtb, region=="Asia", drop=TRUE)
asiamedtb$region <- factor(asiamedtb$region)
asiamedtb$COB <- factor(asiamedtb$COB)

asialowtb <- subset(lowtb, region=="Asia", drop=TRUE)
asialowtb$region <- factor(asialowtb$region)
asialowtb$COB <- factor(asialowtb$COB)
```

####Plots of time to status adjustment by visa types for status adjusters from high TB incidence asian countries

```{r "Asian high TB incidence countries",fig.height=8,fig.width=15,warning=FALSE}
xref=data.frame(xval=c(2.5, 5))
yref=data.frame(yval=c(0.5, 0.8))

#China
china <- subset(asiahightb, COB =="China", drop=TRUE) 
china$COB <- factor(china$COB)
china$visacat.yecai <- factor(china$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=china)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)
 
p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from China \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(china, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df

#India
india <- subset(asiahightb, COB =="India", drop=TRUE) 
india$COB <- factor(india$COB)
india$visacat.yecai <- factor(india$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=india)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)
 
p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from India \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(india, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df

#Pakistan
pakistan <- subset(asiahightb, COB =="Pakistan", drop=TRUE) 
pakistan$COB <- factor(pakistan$COB)
pakistan$visacat.yecai <- factor(pakistan$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=pakistan)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)

p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from Pakistan \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(pakistan, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df


#Philippines
philippines <- subset(asiahightb, COB =="Philippines", drop=TRUE) 
philippines$COB <- factor(philippines$COB)
philippines$visacat.yecai <- factor(philippines$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=philippines)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)

p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from Philippines \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(philippines, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df

#Thailand
thailand <- subset(asiahightb, COB =="Thailand", drop=TRUE) 
thailand$COB <- factor(thailand$COB)
thailand$visacat.yecai <- factor(thailand$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=thailand)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)

p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from Thailand \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(thailand, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df

#Vietnam
vietnam <- subset(asiahightb, COB =="Vietnam", drop=TRUE) 
vietnam$COB <- factor(vietnam$COB)
vietnam$visacat.yecai <- factor(vietnam$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=vietnam)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)

p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from Vietnam \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(vietnam, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df
```

####Plots of time to status adjustment by visa types for status adjusters from medium TB incidence asian countries

```{r "Asian medium TB incidence countries",fig.height=8,fig.width=15,warning=FALSE}
xref=data.frame(xval=c(2.5, 5))
yref=data.frame(yval=c(0.5, 0.8))

#Iran
iran <- subset(asiamedtb, COB =="Iran", drop=TRUE) 
iran$COB <- factor(iran$COB)
iran$visacat.yecai <- factor(iran$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=iran)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)
 
p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from Iran \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(iran, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df

#Japan
japan <- subset(asiamedtb, COB =="Japan", drop=TRUE) 
japan$COB <- factor(japan$COB)
japan$visacat.yecai <- factor(japan$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=japan)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)
 
p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from Japan \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(japan, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df

#South Korea
skorea <- subset(asiamedtb, COB =="S Korea", drop=TRUE) 
skorea$COB <- factor(skorea$COB)
skorea$visacat.yecai <- factor(skorea$visacat.yecai)

##Strata - by visa types
t.survfit <- survfit(Surv(texamyr, status)~visacat.yecai, data=skorea)
t.survfit$cumevent <- 1- t.survfit$surv # to create cumulative frequency variable (1-surv)
t.survframe <- createSurvivalFrame(t.survfit)
 
p <- qplot_survival(t.survframe, FALSE) + xlab("time to exam (years)") + ylab("Cumulative probability of adjusting status") + opts(title="Time to status adjustment in non-immigrant applicants from South Korea \nby Visa Type") + opts(legend.justification=c(1,0), legend.position=c(1,0))

p + geom_vline(aes(xintercept=xval), data=xref, color="blue", linetype="dashed") + geom_hline(aes(yintercept=yval), data=yref, color="blue", linetype="dashed") + scale_y_continuous(labels=percent) + scale_x_continuous(limits=c(0, 20), breaks=c(0, 2.5, 5, 7.5, 10, 15, 20))

##Summary
df <- ddply(skorea, .(visacat.yecai), function(x) c(count=nrow(x), min=min(x$texamyr), mean=mean(x$texamyr), max=max(x$texamyr), quantile=quantile(x$texamyr, seq(.25, .75, by=.25))))

df <- subset(df, visacat.yecai!="NA", drop=TRUE)

df <- rename(df, c(visacat.yecai="Visa type (Yecai's classification)", count="Number", mean="Mean time to exam", min="Min.time to exam", max="Max.time to exam"))
df <- df[order(-df$Number),]
df
```





